---
title: 'Preparing the MEF ATAC-seq data for Etv2 peaks at D1 post induction
output: html_document
date: '12/14/2019'
author: 'Wuming Gong'
---

```{r, message = FALSE}
library(roxygen2)
library(devtools)
#devtools::document('analysis/seatac/packages/seatac')
devtools::load_all('packages/compbio')
devtools::load_all('analysis/seatac/packages/seatac')
library(futile.logger); flog.threshold(TRACE)
library(BSgenome.Mmusculus.UCSC.mm10)
library(Rtsne)
```


# Prepare the windows
```{r}
window_size <- 320
bin_size <- 5
genome <- 'mm10'
bed_file <- '/panfs/roc/scratch/gongx030/datasets/dataset=Etv2PioneerChIPseq_version=20191203a/MEF_Dox_d1_Etv2_summits.bed'
mnase_file <- '/panfs/roc/scratch/gongx030/datasets/dataset=Chronis_version=20170519a/MNase_treat_pileup.bw'
bam_file <- '/panfs/roc/scratch/gongx030/datasets/dataset=Etv2ATAC_version=20190228b/MEF_NoDox.bam'
blacklist_file <- '/panfs/roc/scratch/gongx030/datasets/datasets=blacklists_version=20190827a/mm10.blacklist.bed.gz'

x <- read.table(bed_file, sep = '\t',  header = FALSE)
peaks <- GRanges(seqnames = x[, 1], range = IRanges(x[, 2], x[, 3]))
#set.seed(1); peaks <- sample(peaks, 100)

devtools::load_all('analysis/seatac/packages/seatac'); windows <- prepare_windows(peaks, window_size = window_size, bin_size = bin_size, bigwig_files = c(MEF_MNase = mnase_file), bam_files = bam_file, blacklist_file = blacklist_file, genome = BSgenome.Mmusculus.UCSC.mm10)
saveRDS(windows, file = sprintf('%s/MEF_Dox_d1_Etv2_on_MEF_ATAC.rds', project_dir('seatac')))

```

```{r}
windows <- readRDS(sprintf('%s/MEF_Dox_d1_Etv2_on_MEF_ATAC.rds', project_dir('seatac')))
set.seed(1); gr <- sample(windows[windows$num_reads >= 20], 5000)
latent_dim <- 2; epochs <- 50
set.seed(1); devtools::load_all('analysis/seatac/packages/seatac'); model <- seatac(gr, type = 'vae_20191216c', latent_dim = latent_dim, epochs = epochs)

devtools::load_all('analysis/seatac/packages/seatac'); gr <- model %>% predict(gr)


k <- 2
#cls <- kmeans(gr$latent, k)$cluster
cls <- kmeans(model$data$fragment_size, k)$cluster
#cls2 <- kmeans(model$data$position, k)$cluster

par(mfrow = c(4 + 1, k))
lapply(1:k, function(i) image(matrix(colMeans(mcols(gr)$counts[cls == i, ]) - colMeans(mcols(gr)$counts), metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals), col = colorpanel(100, low = 'blue', mid = 'white', high = 'red')))
lapply(1:k, function(i) image(matrix(colMeans(mcols(gr)$counts[cls == i, ]), metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals), col = colorpanel(100, low = 'blue', mid = 'white', high = 'red')))
lapply(1:k, function(i) plot(colMeans(model$data$fragment_size[cls == i, ])))
lapply(1:k, function(i) plot(colMeans(model$data$position[cls == i, ])))
smoothScatter(gr$latent, main = 'v-plot')
image(matrix(colMeans(mcols(gr)$counts), metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals), col = colorpanel(100, low = 'blue', mid = 'white', high = 'red'))


plot(y, main = 'v-plot')


plot(y2, main = 'v-plot + fragment_size')
y2 <- Rtsne(gr$h_fragment_size)$Y
plot(y2, main = 'fragment_size')

cls <- kmeans(gr$h_fragment_size, 3)$cluster
lapply(1:3, function(i) plot(colMeans(model$data$fragment_size[cls == i, ])))

cls <- kmeans(cbind(gr$latent, gr$h_fragment_size), 3)$cluster
lapply(1:3, function(i) plot(colMeans(model$data$fragment_size[cls == i, ])))


```

source('analysis/seatac/windows.r'); save_windows(windows, gs, window_size, bin_size, fs[1:2], fs[3])

# --- load training windows
source('analysis/seatac/windows.r'); windows <- load_windows(gs, window_size, bin_size, fs[1:2], fs[3])

# --- train model
latent_dim <- 2; epochs <- 10; bs <- 256; spe <- 20
source('analysis/seatac/model_dir_name.r'); model_dir <- model_dir_name(gs, latent_dim, window_size, bin_size, fs[1:2], fs[3], epochs, bs, spe)
windows <- windows[windows$num_reads >= 20]
devtools::load_all('analysis/seatac/packages/seatac'); model <- seatac(windows, latent_dim = latent_dim, epochs = epochs, batch_size = bs, steps_per_epoch = spe)
devtools::load_all('analysis/seatac/packages/seatac'); gr <- model %>% predict(windows)



par(mfcol = c(3, 5))
k <- 5
set.seed(1); cls <- kmeans(gr$latent, k)$cluster
res <- lapply(1:k, function(c){
	i <- cls == c
	xx <- as.matrix(colMeans(mcols(gr)$counts[i, , drop = FALSE]))
	X <- matrix(xx, metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals)
	image(X)
	abline(v = 0.5)
})

smoothScatter(gr$latent, xlim = c(-10, 10), ylim = c(-10, 10))

z <- c(-5, -5)
xx <- model$decoder(matrix(z, nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t()
plot(rowMeans(xx))
plot(colMeans(xx))
col <- colorpanel(100, low = 'blue', mid = 'white', high = 'yellow')
breaks <- seq(0, 1, length.out = 101)
image(model$decoder(matrix(z, nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t(), col = col, breaks = breaks)

par(mfrow = c(3, 4))
hs <- 1:12 + 1000
res <- lapply(hs, function(i){ 
	X <- as(matrix(mcols(gr)$counts[i, ], metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals), 'dgCMatrix')
	xx <- as.matrix(summary(X))[, 1:2]
	x <- cbind(dist_to_center = abs(xx[, 1] -  metadata(gr)$n_bins_per_window / 2), fragment_size = xx[, 2], group = i)
#	plot(x[, 'dist_to_center'] + 1, (x[, 'dist_to_center'] + 1) / x[, 'fragment_size'], log = 'xy', xlim = c(1, metadata(gr)$n_bins_per_window / 2), ylim = c(1 / metadata(gr)$n_intervals, metadata(gr)$n_bins_per_window / 2))
	plot(x[, 'dist_to_center'] + 1, x[, 'fragment_size'] / (x[, 'dist_to_center'] + 1), log = 'xy', xlim = c(1, metadata(gr)$n_bins_per_window / 2), ylim = c(1 / (metadata(gr)$n_bins_per_window / 2 + 1), metadata(gr)$n_intervals))
})

smoothScatter(x[, 1:2])

par(mfrow = c(3, 4))
image(model$decoder(matrix(c(-5, 0), nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t())
image(model$decoder(matrix(c(5, 0), nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t())

image(model$decoder(matrix(c(0, -5), nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t(), col = col, breaks = breaks)
image(model$decoder(matrix(c(5, 0), nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t(), col = col, breaks = breaks)
image(model$decoder(matrix(c(-5, 0), nrow = 1, ncol = 2))$mean() %>% tf$squeeze() %>% as.matrix() %>% t(), col = col, breaks = breaks)

xx <- as.matrix(colMeans(mcols(gr)$counts))
X <- matrix(xx, metadata(gr)$n_bins_per_window, metadata(gr)$n_intervals)
bg_feature <- colMeans(X)
bg_input <- rowMeans(X)
X <- t(t(X) - bg_feature) - bg_input
image(X)

